---
title: "Lab2_Data_Visualization"
format: html
editor: visual
---

# Data Visualization

## Load data from s3

```{r}
install.packages("aws.s3")
install.packages("readr")
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(purrr)
library(aws.s3)
library(readr)

Sys.setenv("AWS_ACCESS_KEY_ID" = "redacted",
           "AWS_SECRET_ACCESS_KEY" = "redacted",
           "AWS_DEFAULT_REGION" = "us-east-1")

bucket <- "tweetsentimentdata"
```

## Load train dataset

```{r}
file_name <- "Complete_Train_data_filtered.csv"
obj <- get_object(object = file_name, bucket = bucket)
tweets <- read_csv(rawToChar(obj))
```

## Load test dataset

```{r}
file_name2 <- "Complete_Test_data_filtered.csv"
obj2 <- get_object(object = file_name2, bucket = bucket)
test_tweets <- read_csv(rawToChar(obj2))
```

## Data processing

```{r}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(purrr)
library(readr)
library(patchwork)


tweets <- tweets %>% 
  mutate(followers = follower_count) %>% 
  mutate(engagement = engagement_score) %>% 
  mutate(sentiment_label2d = sentiment_2dim_label) %>% 
  mutate(sentiment_score2d = sentiment_2dim_score) %>%
  mutate(sentiment_score1d = sentiment_1dim) %>%
  mutate(topic_ = case_when(
    topic_new == "politics" ~ "politics",
    TRUE ~ "other"
  ))

# filter out tweets with top 2% engagement scores
engagement_98 <- quantile(tweets$engagement, 0.98, na.rm = TRUE)

tweets_filtered <- tweets %>%
  filter(engagement < engagement_98)

str(tweets)
```

#### High-level view of variable distribution

```{r, fig.width=12, fig.height=8}
tweets %>%
  select(word_count, followers, textstat_readability, sentiment_score1d, sentiment_score2d, sentiment_score2d) %>%  # select only numeric columns
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free") +
  theme_minimal()
```

### Investigating Engagement Score

-   compare raw vs log transformed before removing any outliers
-   compare raw vs log transformed after removing 2% outliers

```{r}
# before filtering outliers
plot_engagement <- tweets %>% 
  ggplot(aes(x = engagement)) +
  geom_histogram(aes(y = after_stat(count / sum(count))), bins = 80, fill = "#999999", color = "white") +
  labs(
    x = "Engagement (raw)",
    y = "Proportion",
    title = "Distribution of Engagement"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Log-transformed engagement
plot_engagement_log <- tweets %>% 
  ggplot(aes(x = log(engagement))) +
  geom_histogram(aes(y = after_stat(count / sum(count))), bins = 80, fill = "#999999", color = "white") +
  labs(
    x = "Engagement (log-transformed)",
    y = "Proportion",
    title = "Distribution of Log-Engagement"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Combine plots side by side
plot_engagement | plot_engagement_log
```

Conclusion: log transformation is effective in removing fat tails in engagement scores

```{r}
# plot distribution of engagement score to determine whether/what level to remove outliers
# experimented 2% and 5%
tweets %>% 
  ggplot(aes(x = engagement)) +
  geom_histogram(aes(y = after_stat(count / sum(count))), bins = 100, fill = "#999999", color = "white") +
  geom_vline(aes(xintercept = engagement_98, color = "98th Percentile"), 
             linetype = "dashed", linewidth = 1) +
  scale_color_manual(name = "Reference Line", values = c("98th Percentile" = "red")) +
  labs(x = "Engagement", y = "Proportion", title = "Distribution of Engagement Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Conclusion: 2% handles the extreme outliers

```{r}
# after filtering out 2% outliers
plot_engagement <- tweets_filtered %>% 
  ggplot(aes(x = engagement)) +
  geom_histogram(aes(y = after_stat(count / sum(count))), bins = 80, fill = "#999999", color = "white") +
  labs(
    x = "Engagement (raw)",
    y = "Proportion",
    title = "Distribution of Engagement (removed extreme outliers)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 10)  # adjust size here
  )

# Log-transformed engagement
plot_engagement_log <- tweets_filtered %>% 
  ggplot(aes(x = log(engagement))) +
  geom_histogram(aes(y = after_stat(count / sum(count))), bins = 80, fill = "#999999", color = "white") +
  labs(
    x = "Engagement (log-transformed)",
    y = "Proportion",
    title = "Distribution of Log-Engagement (removed extreme outliers)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 10)  # adjust size here
  )

# Combine plots side by side
plot_engagement | plot_engagement_log
```

**Conclusion:**

After comparing histogram plots, we found log transformation is important in removing the fat tails in engagement scores. Also removing top 2% outliers will help us to avoid potential skew from viral tweets.

### Investigating Sentiment distribution

```{r}
tweets %>% 
  ggplot(aes(x = sentiment_score2d, fill = sentiment_label2d)) +
  geom_histogram(bins = 20, position = "dodge") +
  labs(x = "Sentiment Score", y = "Count", fill = "Sentiment Label", title="Distribution of Sentiment Score (2 Dimensional)") +
  scale_fill_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Follower Count distribution

```{r}
tweets %>% 
  ggplot(aes(x = followers)) +
  geom_histogram() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Topic Distribution

```{r}
tweets %>% 
  ggplot(aes(x = topic_new)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

tweets %>%
  ggplot(aes(x = engagement, y =topic_new, color=sentiment_label2d )) +
  geom_point() +
  labs(title = "Topic vs Engagement",
       x = "Topic",
       y = "Engagement Score") +
  theme_minimal()
```

Low topic score

```{r}
tweets %>% 
  ggplot() + 
  aes(x = topic_score, color = topic_, binwidth = 1) + 
  geom_histogram() 
```

#### Word count Distribution

```{r}
tweets %>% 
  ggplot(aes(x = word_count)) +
  geom_histogram() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Investigating the relationship between sentiment score and our possible Y-variables (engagement, likes, retweets, replies and quotes)

Engagement = likes+retweets+replies+quotes

### Sentiment score vs Engagement

Relationship between sentiment score and engagement for negative/positive (Zoomed In)

```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)

plot1 <- tweets %>%
  ggplot(aes(x = sentiment_score2d, y = engagement, color = sentiment_label2d)) +
  geom_point(alpha = 0.5, size = 1) +
  geom_smooth() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Zoomed out",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Sentiment Label"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  )

plot2 <- tweets %>% 
  ggplot() + 
  aes(x = sentiment_score2d, y = engagement, color = sentiment_label2d) + 
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1000)) +  # zooms in
  geom_point(alpha = 0.25, size = 1) + 
  geom_smooth() +
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Sentiment Label"
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  )

plot1 | plot2
```

### Sentiment score vs likes

Relationship between sentiment score and likes for negative/positive (Zoomed In)

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_score2d, y = likes, color = sentiment_label2d) + 
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 700)) +  # zooms in
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Likes vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Likes",
    color = "Sentiment Label"
  ) 
```

### Sentiment score vs retweets

Relationship between sentiment score and retweets for negative/positive (Zoomed In)

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_score2d, y = retweets,  color = sentiment_label2d) + 
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 150)) +  # zooms in
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Retweets vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Retweets",
    color = "Sentiment Label"
  ) 
```

### Sentiment score vs replies

Relationship between sentiment score and replies for negative/positive (Zoomed in)

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_score2d, y = replies, color = sentiment_label2d) + 
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 100)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Replies vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Replies",
    color = "Sentiment Label"
  ) 
```

### Sentiment score vs quotes

Relationship bettwen sentiment scres and qutoes

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_score2d, y = quotes, color = sentiment_label2d) + 
  coord_cartesian(xlim = c(0, 1), ylim = c(0,50)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Quotes vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Quotes",
    color = "Sentiment Label"
  ) 
```

## Investigating Control features

### Followers vs Engagement

Relationship between follower and engagement for negative/positive (Zoomed In)

```{r}
tweets %>% 
  ggplot() + 
  aes(x = followers, y = engagement, color = sentiment_label2d) + 
  coord_cartesian(xlim = c(0, 100000), ylim = c(0, 500)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Followers",
    subtitle = "Zoomed In",
    x = "Followers",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

### Hashtag count vs Engagement

I don't think we should use this because there is high varience with high hashtag count.

```{r}
tweets %>% 
  ggplot() + 
  aes(x = hashtag_count, y = engagement, color = sentiment_label2d) + 
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 2000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() +
  geom_smooth()  + 
  labs(
    title = "Engagement vs. Hashtag Count",
    subtitle = "Zoomed In",
    x = "Hashtag Count",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

#### Hashtag bool

Instead we can use a bool just to mark if a tweet has any hashtags or not.

```{r}
library(dplyr)

summary_df <- tweets %>%
  group_by(sentiment_label2d, hashtag_bool) %>%
  summarise(mean_engagement = mean(engagement, na.rm = TRUE), .groups = "drop")

ggplot(summary_df, aes(x = factor(hashtag_bool), y = mean_engagement,
                       color = sentiment_label2d, group = sentiment_label2d)) +
  geom_line(aes(group = sentiment_label2d), linewidth = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = c(
    "positive" = "#002676",
    "negative" = "#ffc000",
    "neutral" = "#7f7f7f"  # Add fallback if neutral exists
  )) +
  theme_minimal() +
  labs(
    x = "Hashtag Present",
    y = "Average Engagement",
    title = "Average Engagement by Hashtag Presence and Sentiment"
  )
```

### Word Count vs engagement

```{r}
tweets %>% 
  ggplot() + 
  aes(x = word_count, y = engagement, color = sentiment_label2d ) + 
  coord_cartesian(xlim = c(0, 60), ylim = c(0, 1000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Word Count",
    subtitle = "Zoomed In",
    x = "Word Count",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

### Readability

```{r}
tweets %>% 
  # filter(textstat_readability <30) %>%
  # filter(textstat_readability >-5) %>%
  ggplot() + 
  aes(x = textstat_readability, y = engagement, color = sentiment_label2d ) + 
  coord_cartesian(xlim = c(-5, 30), ylim = c(0, 1000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Readability",
    subtitle = "Zoomed In",
    x = "Readability",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

### Relationship between word count and readability

To see if there is a relationship.

```{r}
tweets %>% 
  ggplot() + 
  aes(x = textstat_readability, y = word_count, color = sentiment_label2d ) + 
  coord_cartesian(xlim = c(-5, 30), ylim = c(0, 60)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Word Count vs. Readability",
    subtitle = "Zoomed In",
    x = "Readability",
    y = "Word Count",
    color = "Sentiment Label"
  ) 
```

### Mentions

Don't include. There doesn't seem to be a meaningful relationship here.

```{r}
tweets %>% 
  ggplot() + 
  aes(x = mention_count, y = engagement, color = sentiment_label2d ) + 
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 5000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Mentions",
    subtitle = "Zoomed In",
    x = "Mentions",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

### Emoji count

Don't include. There doesn't seem to be a meaningful relationship here.

```{r}
tweets %>% 
  ggplot() + 
  aes(x = emoji_count, y = engagement, color = sentiment_label2d ) + 
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 5000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("positive" = "#002676", "negative" = "#ffc000")) +
  geom_point() + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Emoji Count",
    subtitle = "Zoomed In",
    x = "Emoji Count",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

## Topic Investigation

What topics should we include? Should we interact with sentiment label and or sentiment score?

### Negative topics

```{r}
tweets %>% 
  filter(sentiment_label2d == "negative") %>% 
  ggplot() + 
  aes(x = sentiment_label2d , y = engagement, color = topic_) + 
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 2000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  geom_point(alpha = 0.25, size = 1) + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Negative Only",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Sentiment Label"
  ) 
```

### Positive Topic

```{r}
tweets %>% 
  filter(sentiment_label2d == "positive") %>% 
  ggplot() + 
  aes(x = sentiment_label2d , y = engagement, color = topic_) + 
  theme_minimal() +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1000)) +  # zooms in
  geom_point(alpha = 0.25, size = 1) + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Positive Only",
    x = "Sentiment Score",
    y = "Engagement",
    col = "Topic"
  ) 
```

## 1-Dimensional Plot

### Engagement vs sentiment score

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_1dim, y = engagement, color = topic_) + 
  coord_cartesian(xlim = c(-1, 1), ylim = c(0, 1000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  geom_point(alpha = 0.25, size = 1) + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Topic"
  ) 
# +
#   theme(
#     legend.position = "bottom",
#     legend.direction = "horizontal"
#   )
```

### Retweets vs Sentiment Score

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_1dim, y = retweets, color = topic_) + 
  coord_cartesian(xlim = c(-1, 1), ylim = c(0, 200)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  geom_point(alpha = 0.25, size = 1) + 
  geom_smooth() + 
  labs(
    title = "Retweets vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Topic"
  ) 
```

### Zoomed out

```{r}
tweets %>% 
  ggplot() + 
  aes(x = sentiment_1dim, y = engagement, color = topic_) + 
  # coord_cartesian(xlim = c(-1, 1), ylim = c(0, 1000)) +  # zooms in
  theme_minimal() +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  geom_point(alpha = 0.4, size = 1) + 
  geom_smooth() + 
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Zoomed Out",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Topic"
  ) 
```

### Data with median smoothing fit

```{r}
library(dplyr)
library(ggplot2)

# Create bins of sentiment_1dim
tweets_binned <- tweets %>%
  mutate(bin = cut(sentiment_1dim, breaks = seq(-1, 1, by = 0.05))) %>%
  group_by(bin, topic_) %>%
  summarize(
    sentiment_mid = median(sentiment_1dim, na.rm = TRUE),
    median_engagement = median(engagement, na.rm = TRUE),
    .groups = "drop"
  )

# Plot with median line instead of geom_smooth
ggplot(tweets, aes(x = sentiment_1dim, y = engagement, color = topic_)) +
  geom_point(alpha = 0.25, size = 1) +
  geom_line(data = tweets_binned, aes(x = sentiment_mid, y = median_engagement), size = 1) +
  coord_cartesian(xlim = c(-1, 1), ylim = c(0, 1000)) +
  theme_minimal() +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  labs(
    title = "Engagement vs. Sentiment Score",
    subtitle = "Zoomed In",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Topic"
  )
```
