---
title: "Lab2_FinalModel"
format: html
editor: visual
---

## Install packages if needed

```{r}
install.packages("aws.s3")
install.packages("readr")
install.packages("GGally")
install.packages("moments")
```

## Pull Train Data from s3

```{r}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(purrr)
library(aws.s3)
library(readr)

Sys.setenv("AWS_ACCESS_KEY_ID" = "redacted",
           "AWS_SECRET_ACCESS_KEY" = "redacted",
           "AWS_DEFAULT_REGION" = "us-east-1")

bucket <- "tweetsentimentdata"
file_name <- "Complete_Train_data_filtered.csv"

# Load the CSV file directly into a dataframe
obj <- get_object(object = file_name, bucket = bucket)
tweets <- read_csv(rawToChar(obj))

# View the data
head(tweets)
```

## Pull Test Data from s3

```{r}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(purrr)
library(aws.s3)
library(readr)

Sys.setenv("AWS_ACCESS_KEY_ID" = "redacted",
           "AWS_SECRET_ACCESS_KEY" = "redacted",
           "AWS_DEFAULT_REGION" = "us-east-1")

bucket <- "tweetsentimentdata"
file_name <- "Complete_Test_data_filtered.csv"

# Load the CSV file directly into a dataframe
obj <- get_object(object = file_name, bucket = bucket)
tweets <- read_csv(rawToChar(obj))

# View the data
head(tweets)
```

## Data processing for modeling

```{r}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(purrr)
library(readr)


tweets_filtered <- tweets %>% 
  mutate(followers = follower_count) %>% 
  mutate(engagement = engagement_score) %>% 
  mutate(sentiment_ = sentiment_2dim_label) %>% #shorten for stargazer 
  mutate(sentiment_label = sentiment_2dim_label) %>% 
  mutate(sentiment_score = sentiment_2dim_score) %>% 
  mutate(topic_ = case_when(
    topic_new == "politics" ~ "politics",
    TRUE ~ "other"
  ))

# # filter out tweets with top 2% engagement scores
# engagement_98 <- quantile(tweets$engagement, 0.98, na.rm = TRUE)
# 
# tweets_filtered <- tweets %>%
#   filter(engagement < engagement_98)

head(tweets_filtered)
```

## Final Model (4 model version)

```{r}
library(stargazer)
library(sandwich)

tweets_negative_politics <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "politics")

tweets_positive_politics <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "politics")

tweets_negative_other <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "other")

tweets_positive_other <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "other")

neg_politics = lm(log(engagement) ~ sentiment_score + followers + word_count
             , data = tweets_negative_politics)

n_pol = neg_politics # shorten for stargazer 

pos_politics = lm(log(engagement) ~ sentiment_score + followers + word_count
             , data = tweets_positive_politics)

p_pol = pos_politics# shorten for stargazer 

neg_other = lm(log(engagement) ~ sentiment_score + followers + word_count
             , data = tweets_negative_other)

n_ot = neg_other # shorten for stargazer 

pos_other = lm(log(engagement) ~ sentiment_score + followers + word_count
             , data = tweets_positive_other)

p_ot = pos_other

# Compute robust standard errors for each model
np <- vcovHC(neg_politics, type = "HC3")
pp <- vcovHC(pos_politics, type = "HC3")
no <- vcovHC(neg_other, type = "HC3")
po <- vcovHC(pos_other, type = "HC3")

# Use stargazer to display the regression results with robust standard errors
stargazer::stargazer(
  n_pol, 
  p_pol,
  n_ot, 
  p_ot,
  type = "text", 
  column.labels = c("Negative Politics", "Positive Politics", "Negative Other", "Positive Other"),
  model.names = FALSE,
  vcov = list(np, 
              pp,
              no, 
              po)
)
```

### Plot the model on top of the data

```{r}
# Generate a sequence of sentiment scores
sentiment_seq <- seq(min(tweets_filtered$sentiment_score, na.rm = TRUE),
                     max(tweets_filtered$sentiment_score, na.rm = TRUE),
                     length.out = 100)

# Negative politics 
# Create a new data frame for prediction for negative politics 
neg_politics_data <- data.frame(sentiment_score = sentiment_seq, followers = mean(tweets_filtered$followers), word_count = round(mean(tweets_filtered$word_count)))

# Calculate residual standard error
sigma <- summary(neg_politics)$sigma

# Predict log(engagement) and exponentiate to get engagement with adjustment for bias: exp((sigma^2) / 2)
neg_politics_data$engagement <- exp(predict(neg_politics, newdata = neg_politics_data)) * exp((sigma^2) / 2)

neg_politics_data$sentiment_score = -neg_politics_data$sentiment_score

# Positive politics 
# Create a new data frame for prediction for positive politics
pos_politics_data <- data.frame(sentiment_score = sentiment_seq, followers = mean(tweets_filtered$followers), word_count = round(mean(tweets_filtered$word_count)))

# Calculate residual standard error
sigma <- summary(pos_politics)$sigma

# Predict log(engagement) and exponentiate to get engagement 
pos_politics_data$engagement <- exp(predict(pos_politics, newdata = pos_politics_data)) * exp((sigma^2) / 2)

# Negative other 
# Create a new data frame for prediction for negative politics 
neg_other_data <- data.frame(sentiment_score = sentiment_seq, followers = mean(tweets_filtered$followers), word_count = round(mean(tweets_filtered$word_count)))

# Calculate residual standard error
sigma <- summary(neg_other)$sigma

# Predict log(engagement) and exponentiate to get engagement
neg_other_data$engagement <- exp(predict(neg_other, newdata = neg_other_data)) * exp((sigma^2) / 2)

neg_other_data$sentiment_score = -neg_other_data$sentiment_score

# Positive other 
# Create a new data frame for prediction for positive politics
pos_other_data <- data.frame(sentiment_score = sentiment_seq, followers = mean(tweets_filtered$followers), word_count = round(mean(tweets_filtered$word_count)))

# Calculate residual standard error
sigma <- summary(pos_other)$sigma

# Predict log(engagement) and exponentiate to get engagement 
pos_other_data$engagement <- exp(predict(pos_other, newdata = pos_other_data)) * exp((sigma^2) / 2)


# Plot
ggplot(tweets_filtered, aes(x = sentiment_1dim, y = engagement, color = topic_)) +
  geom_point(alpha = 0.25, size = 1) +
  geom_line(data = neg_other_data, aes(x = sentiment_score, y = engagement), color = "#6688BB", size = 1) +
  geom_line(data = pos_other_data, aes(x = sentiment_score, y = engagement), color = "#6688BB", size = 1) +
  geom_line(data = neg_politics_data, aes(x = sentiment_score, y = engagement), color = "#ffc000", size = 1) +
  geom_line(data = pos_politics_data, aes(x = sentiment_score, y = engagement), color = "#ffc000", size = 1) +
  coord_cartesian(xlim = c(-1, 1), ylim = c(0, 1000)) +
  scale_color_manual(values = c("other" = "#002676", "politics" = "#ffc000")) +
  theme_minimal() +
  labs(
    title = "Linear Regression",
    subtitle = "For average followers and word count",
    x = "Sentiment Score",
    y = "Engagement",
    color = "Topic"
  )


```

### Normally distributed residuals? Yes

```{r}
residuals <- c(resid(neg_politics), resid(pos_politics), resid(neg_other), resid(pos_other))
hist(residuals, breaks = 30, main = "Histogram of Residuals: with log(engagement)", xlab = "Residuals")
```

### QQ plot

```{r}
qqnorm(residuals)
qqline(residuals, col = "red", lwd = 2)
```

### Residuals vs Fitted values

```{r}
residuals<- c(resid(neg_politics), resid(pos_politics), resid(neg_other), resid(pos_other))
fitted_vals <- c(neg_politics$fitted.values,  pos_politics$fitted.values, neg_other$fitted.values, pos_other$fitted.values)

df_resid <- data.frame(
  fitted_vals = fitted_vals,
  resid = residuals
)

ggplot(df_resid, aes(x = fitted_vals , y = resid)) +
  geom_point() +
  labs(title = "Residuals vs Fitted values",
       x = "Fitted values",
       y = "Residuals") +
  theme_minimal()
```

## Final model (Single Model version)

```{r}
tweets_filtered$topic_ <- factor(tweets_filtered$topic_)
tweets_filtered$topic_ <- relevel(tweets_filtered$topic_, ref = "politics")

tweets_filtered$sentiment_label <- factor(tweets_filtered$sentiment_label)
tweets_filtered$sentiment_label <- relevel(tweets_filtered$sentiment_label, ref = "negative")

mod1_log = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + hashtag_bool
             , data = tweets_filtered)


stargazer::stargazer(
  mod1_log,
  type="text")
```

### Normally distributed residuals? Yes

```{r}
residuals <- resid(mod1_log)
hist(residuals, breaks = 30, main = "Histogram of Residuals: with log(engagement)", xlab = "Residuals")
```

### QQ plot

```{r}
qqnorm(residuals)
qqline(residuals, col = "red", lwd = 2)
```

### Residuals vs Fitted values

```{r}
residuals<- resid(mod1_log)
fitted_vals <- mod1_log$fitted.values

df_resid <- data.frame(
  fitted_vals = fitted_vals,
  resid = residuals
)

ggplot(df_resid, aes(x = fitted_vals , y = resid)) +
  geom_point() +
  labs(title = "Residuals vs Fitted values",
       x = "Fitted values",
       y = "Residuals") +
  theme_minimal()
```

### Evaluate whether model assumptions are met

```{r}
# Check if residuals are independent
resid <- mod1_log$residuals
obs_stat <- Box.test(resid, lag =10, type="Ljung-Box")
obs_pval <- obs_stat$p.value
obs_stat
n_perm <- 1000
perm_pval <- numeric(n_perm)

for (i in 1:n_perm) {
  perm_resid <- sample(resid)  # randomly permute residuals
  perm_result <- Box.test(perm_resid, lag =10, type="Ljung-Box")
  perm_pval[i] <- perm_result$p.value
}

hist(perm_pval, breaks = 30, main = "Permutation Test: Residual Independence",
     xlab = "p-values")
abline(v = obs_stat, col = "red", lwd = 2)
```
