---
title: "Lab2_Modeling"
format: revealjs
editor: visual
---

## Install packages if needed

```{r}
install.packages("aws.s3")
install.packages("readr")
install.packages("GGally")
install.packages("moments")
```

## Pull Train Data from s3

```{r}
library(aws.s3)
library(readr)

Sys.setenv("AWS_ACCESS_KEY_ID" = "redacted",
           "AWS_SECRET_ACCESS_KEY" = "redacted",
           "AWS_DEFAULT_REGION" = "us-east-1")

bucket <- "tweetsentimentdata"
file_name <- "Complete_Train_data_filtered.csv"

# Load the CSV file directly into a dataframe
obj <- get_object(object = file_name, bucket = bucket)
tweets <- read_csv(rawToChar(obj))

# View the data
head(tweets)
```

## Pull Test Data from s3

```{r}
library(aws.s3)
library(readr)

Sys.setenv("AWS_ACCESS_KEY_ID" = "redacted",
           "AWS_SECRET_ACCESS_KEY" = "redacted",
           "AWS_DEFAULT_REGION" = "us-east-1")

bucket <- "tweetsentimentdata"
file_name <- "Complete_Test_data_filtered.csv"

# Load the CSV file directly into a dataframe
obj <- get_object(object = file_name, bucket = bucket)
tweets <- read_csv(rawToChar(obj))

# View the data
head(tweets)
```

## Data processing

```{r}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(purrr)
library(readr)
library(stargazer)
library(sandwich)
library(lmtest)
library(moments)
library(zoo)


tweets <- tweets %>% 
  mutate(followers = follower_count) %>% 
  mutate(engagement = engagement_score) %>% 
  mutate(sentiment_label = sentiment_2dim_label) %>% 
  mutate(sentiment_ = sentiment_2dim_label) %>% #shorten for stargazer 
  mutate(sentiment_score = sentiment_2dim_score) %>% 
  mutate(sentiment_label2d = sentiment_2dim_label) %>% 
  mutate(sentiment_score2d = sentiment_2dim_score) %>%
  mutate(sentiment_score1d = sentiment_1dim) %>%
  mutate(hashtag_bool = I(hashtag_count>0)) %>%
  mutate(topic_ = case_when(
    topic_new == "politics" ~ "politics",
    TRUE ~ "other"
  ))


# # filter out tweets with top 2% engagement scores
# engagement_98 <- quantile(tweets$engagement, 0.98, na.rm = TRUE)
# 
# tweets_filtered <- tweets %>%
#   filter(engagement < engagement_98)

tweets_filtered <- tweets

head(tweets_filtered)
```

# Model Building

### Model 1: Engagement vs Follower Count (single X variable)

```{r}
options(width = 200)
# plug in follower_count, find the best polynomial for follower_count
mod1a <- lm(engagement ~ followers, data = tweets)
mod1b <- lm(engagement ~ followers + I(followers^2), data = tweets)
mod1c <- lm(log(engagement) ~ followers, data = tweets)

stargazer::stargazer(
  mod1a,
  mod1b,
  mod1c,
  type = "text",
  column.labels = c("Linear", "Quadratic", "Log Outcome"),
  covariate.labels = c("Followers", "Followers²", "Constant"),
  model.names = FALSE,
  no.space = TRUE,       # tighter vertical spacing
  align = TRUE,          # tighter horizontal alignment
  digits = 2,
  omit.stat = c("f", "ser")
)
```

#### Conclusion on follower_count:

Linear and quadratic (follower-count) show similar results, while log transformation on engagement does significantly improve model results - we will go with linear model for follower count.

### Model 2: Engagement vs Sentiment (single X variable)

Attempted different polynomials with sentiment

```{r}
# base model
mod2b1 <- lm(log(engagement) ~ sentiment_score2d + sentiment_label2d, data = tweets) 

# variation1: base model + sentiment^2
mod2b2 <- lm(log(engagement) ~ I(sentiment_score2d^2) + sentiment_label2d, data = tweets) 

#  variation2: base model + logit(sentiment)
tweets$sentiment_logit <- log(tweets$sentiment_score2d / (1 - tweets$sentiment_score2d))
mod2b3 <- lm(log(engagement) ~ sentiment_logit + sentiment_label2d, data = tweets) 

#  variation3: base model + sentiment_score*label
mod2b4 <- lm(log(engagement) ~ sentiment_score2d * sentiment_label2d, data = tweets)


stargazer::stargazer(
  mod2b1,
  mod2b2,
  mod2b3,
  mod2b4,
  type = "text",
  column.labels = c("Linear", "Quadratic", "Logit", "Interaction"),
  model.names = FALSE,
  no.space = TRUE,       # tighter vertical spacing
  align = TRUE,          # tighter horizontal alignment
  digits = 2,
  omit.stat = c("f", "ser")
)
```

```{r}
# base model
mod2b1 <- lm(log(engagement) ~ sentiment_score2d + sentiment_label2d, data = tweets) 

# variation 4: 1 dimension encoding
mod2b5 <- lm(log(engagement) ~ sentiment_score1d, data = tweets) 

# variation 5: 1 dimension encoding + interaction
mod2b6 <- lm(log(engagement) ~ sentiment_score1d + sentiment_label2d, data = tweets) 

options(width = 200)
stargazer::stargazer(
  mod2b1,
  mod2b5,
  mod2b6,
  type = "text",
  column.labels = c("2d", "1d", "1dlabel"),
  model.names = FALSE,
  # no.space = TRUE,       # tighter vertical spacing
  align = TRUE,          # tighter horizontal alignment
  digits = 2,
  omit.stat = c("f", "ser")
)
```

#### Conclusion:

-   Transformation: Taking a 2nd-order polynomial (mod2b2) and taking logit(mod2b3) of sentiment score don't change the model results by much. Notably by interacting labels with score does provide further information about the relationship.

-   Encoding: The 1D encoding explains more variance in engagement on its own. However, when interacted with sentiment labels, both 1D and 2D encodings perform similarly.

**Conclusion** - we will keep the sentiment score in linear form and interact it with sentiment labels.

Since 1d and 2d perform similarly and we are interested in studying how the strength of sentiments impact engagement, we will keep the 2d encoding.

### Model 3: Engagement vs Topic (single X variable)

```{r}
# `topic` with 5 categories
mod3a <- lm(log(engagement) ~ topic_new, data = tweets)

# `topic` with 2 categories, i.e., politics vs others
mod3b <- lm(log(engagement) ~ topic_, data = tweets)

stargazer::stargazer(
  mod3a,
  mod3b,
  type = "text",
  column.labels = c("5_topic_cat", "2_topic_cat"),
  model.names = FALSE,
  no.space = TRUE,       # tighter vertical spacing
  align = TRUE,          # tighter horizontal alignment
  digits = 2
)

tweets %>% 
  ggplot(aes(x = topic_new)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Conclusion:

In the first model - 'topic with 5 categories' - we can see all the topic categories has a significant relationship with engagement. Interestingly, in contrast to 'politics', all four have a (relatively) negative relationship with engagement. So we decided to compare 'politics' with 'other' for this study.

### Model 4: Engagement vs word_count (single X variable)

```{r}
mod4 <- lm(log(engagement) ~ word_count, data = tweets)

# variation1: base model + word_count^2 
mod4a <- lm(log(engagement) ~ word_count + I(word_count^2), data = tweets)

stargazer::stargazer(
  mod4,
  mod4a,
  type = "text",
  column.labels = c("word_count", "wc_squared"),
  model.names = FALSE,
  no.space = TRUE,       # tighter vertical spacing
  align = TRUE,          # tighter horizontal alignment
  digits = 2
)
```

#### Conclusion:

Quadratic doesn't improve the model by much, we will go with the linear form of word_count.

### Model 5: Engagement vs Sentiment + Follower Count + Topic (Combined variables)

```{r, results='asis'}
mod_c1 <- lm(log(engagement) ~ sentiment_score2d + sentiment_label2d + followers + topic_, data = tweets)

mod_c1a <- lm(log(engagement) ~ sentiment_score2d + sentiment_label2d + followers + topic_ + sentiment_label2d:topic_, data = tweets)

stargazer::stargazer(
  mod_c1,
  mod_c1a,
  type = "text",
  column.labels = c("combined", "combined+interaction"),
  model.names = FALSE,
  no.space = TRUE,       # tighter vertical spacing
  align = TRUE,          # tighter horizontal alignment
  digits = 2
)
```

**Conclusion:**

-   `followers` significantly increases engagement and `positive_sentiment` significantly reduces engagement
-   `politics` itself increases engagement while `topic_Other` decreases engagement
-   When interact `politics` with `sentiment label`, we can see tweets with positive sentiment significantly decreases engagement

### Model 6: Engagement vs Sentiment\*label\*topic + Follower Count + Topic (Combined variables)

```{r}
mod_c2a <- lm(log(engagement) ~ sentiment_score2d*sentiment_label2d*topic_ + followers + word_count, data = tweets)
mod_c2b <- lm(log(engagement) ~ sentiment_score1d*sentiment_label2d*topic_ + followers + word_count, data = tweets)

se1 <- vcovHC(mod_c2a, type = "HC1")
se2 <- vcovHC(mod_c2b, type = "HC1")

stargazer::stargazer(
  mod_c2a,
  mod_c2b,
  se = list(se1, se2),
  type = "text", 
  column.labels = c("Sentiment_2d", "Sentiment_1d"),
  model.names = FALSE,
  # no.space = TRUE,
  align= TRUE,
  omit.stat = c("f", "ser")
)
```

```{r}
# with filtered data, no log
mod_c3a <- lm(engagement ~ sentiment_score2d*sentiment_label2d*topic_ + followers + word_count, data = tweets_filtered)
mod_c3b <- lm(engagement ~ sentiment_score1d*sentiment_label2d*topic_ + followers + word_count, data = tweets_filtered)


se1 <- vcovHC(mod_c3a, type = "HC1")
se2 <- vcovHC(mod_c3b, type = "HC1")

stargazer::stargazer(
  mod_c3a,
  mod_c3b,
  se = list(se1, se2),
  type = "text", 
  column.labels = c("Sentiment_2d", "Sentiment_1d"),
  model.names = FALSE,
  # no.space = TRUE,
  align= TRUE,
  omit.stat = c("f", "ser")
)
```

```{r}
resid <- mod_c2a$residuals
fitted_vals <- mod_c2a$fitted.values

df_resid <- data.frame(
  fitted_vals = fitted_vals,
  resid = resid
)

ggplot(df_resid, aes(x = fitted_vals , y = resid)) +
  geom_point() +
  labs(title = "Residuals vs Fitted values",
       x = "Fitted values",
       y = "Residuals") +
  theme_minimal()
```

```{r}
library(lmtest)
bptest(mod_c2b)
```

### Conclusion

Log of engagement produces a better model here. We can see that the 1D and 2D sentimetn score have the same coefficients but 1D has a negative coefficient for the negative sentiment label.

### Model 7: Engagement vs Sentiment\*label\*topic + Follower Count + Topic + avg_engegement

Average engagement is the average engagement for a user across data from 2009 to 2023.

```{r}
tweets_filtered$topic_new <- factor(tweets_filtered$topic_new)
tweets_filtered$topic_new <- relevel(tweets_filtered$topic_new, ref = "politics")

mod7a = lm(log(engagement) ~ sentiment_score*sentiment_*topic_new + followers 
             , data = tweets_filtered)

mod7b = lm(log(engagement) ~ sentiment_score*sentiment_*topic_new + followers + avg_engagement 
             , data = tweets_filtered)


stargazer::stargazer(
  mod7a, mod7b,
  type="text")

```

### Conclusion

Adding in the users average engagement changes the followers coefficient goes to nearly 0 meaning that average engagement and followers explain some of the same variance of the model. Additionally, there is a concern here that average user engagement corelates with to the topic as well. We decided to remove average user engagement because we want to see the effects of topic and we worry that they will be reduced by adding in this variable.

### Model 8: Adding hashtag_bool, hashtag count, mention_count & readability to see if imporoves the model

```{r}

mod_base = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_filtered)

mod_hash_count = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + hashtag_count
             , data = tweets_filtered)

mod_hash_bool = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + hashtag_bool
             , data = tweets_filtered)

mod_mention = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + mention_count
             , data = tweets_filtered)

mod_readabiliy = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + textstat_readability
             , data = tweets_filtered)


stargazer::stargazer(
  mod_base, mod_hash_count,mod_hash_bool, mod_mention, mod_readabiliy,
  type="text")

```

### Conclusion

All of these additional variables are significant, but none improve the R\^2 significantly and none have a better f-statistic than the original model. We chose to not include them.

## Model 9: Comparing likes, retweets, replies and quotes

```{r}

tweets_filtered$topic_ <- factor(tweets_filtered$topic_)
tweets_filtered$topic_ <- relevel(tweets_filtered$topic_, ref = "politics")

tweets_filtered$sentiment_label <- factor(tweets_filtered$sentiment_label)
tweets_filtered$sentiment_label <- relevel(tweets_filtered$sentiment_label, ref = "negative")

# Create a z-score for each feature that is not catagorical 
tweets_z = tweets_filtered %>% 
  mutate(
    followers = scale(followers)[,1],
    word_count = scale(word_count)[,1],
    engagement = scale(engagement)[,1],
    likes = scale(likes)[,1],
    retweets = scale(retweets)[,1],
    replies = scale(replies)[,1], 
    quotes = scale(quotes)[,1]
  )


lm_log = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_z)

lm_replies = lm(log(replies) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_z)

lm_likes = lm(log(likes) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_z)

lm_retweets = lm(log(retweets) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_z)

lm_replies = lm(log(replies) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_z)

lm_quotes = lm(log(quotes) ~ sentiment_score*sentiment_*topic_ + followers + word_count
             , data = tweets_z)



stargazer::stargazer(
  lm_log, 
  lm_likes,
  lm_retweets,
  lm_replies,
  lm_quotes,
  type="text")
```

## Conclusion

Retweets have the biggest increase in engagement as the strength of sentiment increases. Interestingly this relationship is stronger than with replies and likes. There is not a statistically significant relationship between strength of sentiment and quotes for negative politics.

## Model 10: Removing outliers

```{r}
# filter out tweets with top 2% engagement scores
engagement_98 <- quantile(tweets$engagement, 0.98, na.rm = TRUE)

tweets_filtered$topic_ <- factor(tweets_filtered$topic_)
tweets_filtered$topic_ <- relevel(tweets_filtered$topic_, ref = "politics")

tweets_filtered$sentiment_label <- factor(tweets_filtered$sentiment_label)
tweets_filtered$sentiment_label <- relevel(tweets_filtered$sentiment_label, ref = "negative")

mod1_log = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + hashtag_bool
             , data = tweets_filtered)

mod1_log_no_outliers = lm(log(engagement) ~ sentiment_score*sentiment_*topic_ + followers + word_count + hashtag_bool
             , data = tweets_filtered %>% filter(engagement < engagement_98))

stargazer::stargazer(
  mod1_log,mod1_log_no_outliers, 
  type="text")
```

Residuals vs fitted values with outliers removed

```{r}
library(tidyverse)
library(patchwork)
hist(resid(mod1_log_no_outliers), breaks = 30, main = "Histogram of Residuals: Outliers removed", xlab = "Residuals")
hist(resid(mod1_log), breaks = 30, main = "Histogram of Residuals: Outliers not removed", xlab = "Residuals")


```

## Conclusion

Removing outliers does not help make the residuals more normal.

## Model 11: 4 separate models to ease interpretation, Without log

```{r}
library(stargazer)
library(sandwich)

tweets_negative_politics <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "politics")

tweets_positive_politics <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "politics")

tweets_negative_other <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "other")

tweets_positive_other <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "other")

neg_politics = lm(engagement ~ sentiment_score
             + followers
             + word_count
             , data = tweets_negative_politics)

n_pol = neg_politics


pos_politics = lm(engagement ~ sentiment_score
             + followers
             + word_count
             , data = tweets_positive_politics)

p_pol = pos_politics

neg_other = lm(engagement ~ sentiment_score
             + followers
             + word_count
             , data = tweets_negative_other)

n_ot = neg_other

pos_other = lm(engagement ~ sentiment_score
             + followers
             + word_count
             , data = tweets_positive_other)

p_ot = pos_other

# Compute robust standard errors for each model
np <- vcovHC(neg_politics, type = "HC3")
pp <- vcovHC(pos_politics, type = "HC3")
no <- vcovHC(neg_other, type = "HC3")
po <- vcovHC(pos_other, type = "HC3")

# Use stargazer to display the regression results with robust standard errors
stargazer::stargazer(
  n_pol, 
  p_pol,
  n_ot, 
  p_ot,
  type = "text", 
  column.labels = c("Negative Politics", "Positive Politics", "Negative Other", "Positive Other"),
  model.names = FALSE,
  vcov = list(np, 
              pp,
              no, 
              po)
)
```

Normally distributed residuals? No

```{r}
residuals <- c(resid(neg_politics), resid(pos_politics), resid(neg_other), resid(pos_other))
hist(residuals, breaks = 30, main = "Histogram of Residuals", xlab = "Residuals")
```

## Conclusion

This model is very interpretable. As you go from completely neutral to completely negative for politics you increase engagement by 553 on average, holding everything else constant. This model is not trustworthy though because it has highly skewed residuals. We will use the next model, that is less interpretable, but has normally distributed residuals.

## Model 12: 4 separate models with log

### FINAL MODEL

```{r}
library(stargazer)
library(sandwich)

tweets_negative_politics <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "politics")

tweets_positive_politics <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "politics")

tweets_negative_other <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "other")

tweets_positive_other <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "other")

neg_politics = lm(log(engagement) ~  sentiment_score + followers + word_count
             , data = tweets_negative_politics)

n_pol = neg_politics # shorten for stargazer 

pos_politics = lm(log(engagement) ~ sentiment_score + followers + word_count
             , data = tweets_positive_politics)

p_pol = pos_politics# shorten for stargazer 

neg_other = lm(log(engagement) ~  sentiment_score + followers + word_count
             , data = tweets_negative_other)

n_ot = neg_other # shorten for stargazer 

pos_other = lm(log(engagement) ~ sentiment_score + followers + word_count
             , data = tweets_positive_other)

p_ot = pos_other

# Compute robust standard errors for each model
np <- vcovHC(neg_politics, type = "HC3")
pp <- vcovHC(pos_politics, type = "HC3")
no <- vcovHC(neg_other, type = "HC3")
po <- vcovHC(pos_other, type = "HC3")

# Use stargazer to display the regression results with robust standard errors
stargazer::stargazer(
  n_pol, 
  p_pol,
  n_ot, 
  p_ot,
  type = "text", 
  column.labels = c("Negative Politics", "Positive Politics", "Negative Other", "Positive Other"),
  model.names = FALSE,
  vcov = list(np, 
              pp,
              no, 
              po)
)
```

### Normally distributed residuals? Yes

```{r}
residuals <- c(resid(neg_politics), resid(pos_politics), resid(neg_other), resid(pos_other))
hist(residuals, breaks = 30, main = "Histogram of Residuals: with log(engagement)", xlab = "Residuals")
```

## Conclusion

This is the best and more interpretable model yet.

## Model 12: Same as final model but with 1D sentiment column

```{r}
library(stargazer)
library(sandwich)

tweets_negative_politics <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "politics")

tweets_positive_politics <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "politics")

tweets_negative_other <- tweets_filtered %>% 
  filter(sentiment_label == "negative", 
         topic_ == "other")

tweets_positive_other <- tweets_filtered %>% 
  filter(sentiment_label == "positive", 
         topic_ == "other")

neg_politics = lm(log(engagement) ~ 
              sentiment_1dim
             + followers
             + word_count
             , data = tweets_negative_politics)

n_pol = neg_politics # shorten for stargazer 

pos_politics = lm(log(engagement) ~ 
                  sentiment_1dim
             + followers
             + word_count
             , data = tweets_positive_politics)

p_pol = pos_politics# shorten for stargazer 

neg_other = lm(log(engagement) ~ 
               sentiment_1dim
             + followers
             + word_count
             , data = tweets_negative_other)

n_ot = neg_other # shorten for stargazer 

pos_other = lm(log(engagement) ~ 
               sentiment_1dim
             + followers
             + word_count
             , data = tweets_positive_other)

p_ot = pos_other

# Compute robust standard errors for each model
np <- vcovHC(neg_politics, type = "HC3")
pp <- vcovHC(pos_politics, type = "HC3")
no <- vcovHC(neg_other, type = "HC3")
po <- vcovHC(pos_other, type = "HC3")

# Use stargazer to display the regression results with robust standard errors
stargazer::stargazer(
  n_pol, 
  p_pol,
  n_ot, 
  p_ot,
  type = "text", 
  column.labels = c("Negative Politics", "Positive Politics", "Negative Other", "Positive Other"),
  model.names = FALSE,
  vcov = list(np, 
              pp,
              no, 
              po)
)
```

## Conclusion

This model produces the same results, but with negative slopes for the negative labels which makes it easy to show on a 1D plot (negative sentiment on the right side and postive on the left side).
